---
title: "3_31_2020_Winslow_Thesis_Poisson_Model"
author: "James Winslow"
date: "4/1/2020"
output: pdf_document
---
```{r Import_Data}
source("Winslow_Functions.R")

# this function creates a global object called thesis_data #
# make sure this isn't problematic since using '<<-' isn't a great habit #

import_data("local")
```

```{r}
source("Winslow_Functions.R")
library(dplyr)
library(reshape)

# this function creates a global object called Reshaped & cast_thesis #
# make sure this isn't problematic since using '<<-' isn't a great habit #

clean_data(thesis_data)

```


```{r}
# make a copy of the data for this next bit #
simple_model <- cast_thesis

# rename the variables so it's clear what we are using them for in the simple
# model
simple_model <- `colnames<-`(simple_model, c("id","time.to.death","n.events","dead"))
simple_model$log.time <- log(simple_model$time.to.death)
```
The simplest way to model count data is using the Poisson distribution. The density can be written \\
\begin{math}
f(y|\lambda) = \frac{e^{- \lambda}\lambda^{y}}{y!}
\end{math}

We want to construct a poisson model for the number of infections observed up to a certain time. By imposing an offset, we can interpret the model as a rate instead of a count. This is beneficial to manage the variable start times that subjects may have  

$log(\lambda_{x}) = log(t_{x}) + \beta^{'}_{0} + \beta^{x}_{1}$ \newline
where $log(t_{x})$ is the offset. \\

In a GLM, the canonical link is written as \\
\begin{math}
g(\lambda) = log(\lambda)
\end{math}
This describes a log-linear relationship between the linear predictor and the mean (of the outcome). 
```{r}
# the model we want to use has a poisson component which will link to the survival model via the fraily (random effect) #
# this chunk does some diagnostics to see how well event count fits pois dist #
library(car)
library(MASS)
# use offset()

fish <- glm(n.events ~ offset(log.time), family = poisson(link = "log"),
              data  = simple_model)

summary(fish)
# Coefficients:
#             Estimate Std. Error z value Pr(>|z|)    
# (Intercept) -9.09893    0.02645    -344   <2e-16 ***
```

Due to the Poisson distirbution being a 1-parameter family, the mean and variance are equal. When this is the case, we say that the data is dispersed with some parameter $\Phi = 1$. If we want to relax the assumption that the mean and variance are equal (which is almost certainly the case) we may employ the Quasi-Poisson distribution which allows $\Phi$ to be estimated from the data. 
```{r}
quasi_fish <- glm(n.events ~ offset(log.time), family = quasipoisson(link = "log"),
              data  = simple_model)
summary(quasi_fish)
quasi_fish$coefficients
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept) -9.09893    0.04179  -217.7   <2e-16 ***
```
As we can see, the dispersion parameter here is 2.50. This means that the data is over-dispersed. If we were to find out that adding covariates as predictors to the model doesn't help, we might think that the overdispersion is due to heterogeneity. \\
Also note that the coefficient estimates are equal (this should ALWAYS be the case). The difference is that inference done here would be adjusted for the adjusted standard error(s). \\
There is no fully specified likelihood for the quasi-Poisson model. 
```{r}
time <-seq (0,8154,0.1)  

values <- predict(fish, list(log.time = time))

plot(simple_model$log.time, simple_model$n.events, xlab = "Offset(Days)", ylab = "Events", pch = 16)

lines(time, exp(values), lwd = 2, col = "blue")
```

We may also model over-dispersed data such as ours using the negative binomial distribution. The negative binomial may arise from a Poisson$(\lambda)$ distribution where the parameter $(\lambda)$ follows a Gamma distribution. 
```{r}
# this chunk does some diagnostics to see how well event count fits neg bin dist #
library(car)
library(MASS)
# need to figure out what is happening with the parameterization in R for neg bin

neg_bin_model <- glm.nb(n.events ~ offset(log.time), link = "log", data  = simple_model)
summary(neg_bin_model)
```

```{r}
# Check to see if we want to use coeftest(fish,vcov=sandwich)
library(lmtest)
library(sandwich)
coeftest(fish,vcov=sandwich)
```


Now, we will investigate the impact of including covariate data in our model. We can write the standard log-linear Poisson regression model as \newline

\begin{align*}
log[E(Y_{i}|X_{i})] &= log(\mu_{i}) \\
                      &= log(T_{i}) + \beta_{1}X_{1} + \beta_{2}X_{2} + \cdots + \beta_{i}X_{i}
\end{align*}



### Count Models With Covariates
The covariates available are : hospital site(2 versions), gender, race, organ, quarter, year, lt2
```{r}
# rename the variables so it's clear what we are adding covariates now
head(Reshaped)
cov_model <- `colnames<-`(Reshaped,c("id","time.to.death","n.events","dead","hospital","gender","race","organ","quarter","year","lt2","UngrpHospital"))

head(cov_model)
cov_model$log.time <- log(cov_model$time.to.death)
```

```{r}
# the model we want to use has a poisson component which will link to the survival model via the fraily (random effect) #
# this chunk does some diagnostics to see how well event count fits pois dist #
library(car)
library(MASS)
# 

covariate_fish <- glm(n.events ~ offset(log.time) + hospital + gender + race + organ + year + lt2, family = poisson(link = "log"),
              data  = cov_model)

summary(covariate_fish)

```

We may also model over-dispersed data such as ours using the negative binomial distribution. The negative binomial may arise from a Poisson$(\lambda)$ distribution where the parameter $(\lambda)$ follows a Gamma distribution. 
```{r}
# this chunk does some diagnostics to see how well event count fits neg bin dist #
library(car)
library(MASS)
# need to figure out what is happening with the parameterization in R for neg bin

cov_neg_bin_model <- glm.nb(n.events ~ offset(log.time) + hospital + gender + race + organ + year + lt2, control = glm.control(maxit=100),link = "log", data  = cov_model)
summary(cov_neg_bin_model)
```